PRIMUS: Galaxy Environment on the Quiescent Fraction Evolution at $z < 0.8$

### 2014-10-09 qfenv_paper/qf_environment.tex
Changed H0 = 100 to H0 = 70

Throughout the paper we assume cosmology with $\Omega_{m} = 0.3, \Omega_{\Lambda} = 0.7$, and $H_0 = 70\;h\: \rm{km} \; \rm{s}^{-1} \rm{Mpc}^{-1}$ with $h = 0.7$.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% SAMPLE SELECTION
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Sample Selection} \label{sec:sample}
We are interested in quantifying the effects of galaxy environment on the evolution of the quenched fraction over the redshift range $0 < z < 1$. For our analysis, we require a sample with sufficient depth and robustness to probe the redshift range and to measure galaxy environment. PRIMUS with its $\sim 120,000$ spectroscopic redshifts provides a large data set at intermediate redshifts for our analysis. In addition, we anchor our analysis at intermediate redshifts with a low redshift sample derived from the Sloan Digital Sky Survey (\cite{York:2000aa}).

@@

For our aperture, we use a cylinder of dimensions: $R_{\rm{ap}} = 2 \; h^{-1} \rm{Mpc}$ and $H_{\rm{ap}} = 25 \; h^{-1}\rm{Mpc}$. Though spherical apertures are often used in literature (e.g. \citealt{Croton:2005aa}), we use a cylindrical aperture in order to account for the PRIMUS redshift errors and redshift space distortions (i.e. ``Finger of God" effect). \cite{Cooper:2005aa} finds that $\pm 1000 \; \rm{km} \; \rm{s^{-1}}$ optimally reduces the effects of redshift space distortions. The PRIMUS redshift uncertainty at $z \sim 0.7$ corresponds to $\sigma_z < 0.01$, so our aperture height choice of $25 \; \rm{Mpc}/h$ accounts for both of these effects. Our choice of cylinder radius was motivated by scale dependence analyses in literature (\citealt{Blanton:2006aa, Wilman:2010aa, Muldrew:2012aa}), which suggest that galactic properties such as color and quenched fractions are dependent on halo-scale properties such as host dark matter halo mass. \cite{Wilman:2010aa}, which uses environment defined by annuli of different radii, find positive correlation for quenched fraction and color on scales $< 1 \; \rm{Mpc}$ and anti-correlation on scales $> 3 \; \rm{Mpc}$. Our choice of $2 \rm{Mpc}/h$ provides sufficient sample size of galaxies in dense environments, for robust statistics, while tracing galactic properties within the halo scale.

@@

The analysis we describe below uses a fixed cylindrical aperture with dimensions $R_{\rm{ap}} = 2 \:h^{-1}\rm{Mpc}$ and $H_{\rm{ap}} = 25 \:h^{-1}\rm{Mpc}$ to measure environment. The same analysis was extended for varying aperture dimensions $R_{\rm{ap}} = 1, 2, 3 \: h^{-1}\rm{Mpc}$ and $H_{\rm{ap}} = 25, 50 \:h^{-1}\rm{Mpc}$. We also adjusted the environment classification thresholds for these analyses. The results obtained from using different apertures and environment classification thresholds are qualitatively consistent with the results presented below.

@@

To account for these edge effects, we use a Monte Carlo method to impose edge cutoffs on our galaxy sample. First, using \texttt{ransack} from \cite{Swanson:2008aa}, we construct a random sample of  $N_{\rm{ransack}} = 1,000,000$ points with $RA$ and $Dec$ randomly selected within the window function of the EDP (SDSS EDP and PRIMUS EDP separately). We then compute the angular separation, $\theta_{i, \rm{ap}}$ that corresponds to $R_{\rm{ap}}$ (Section \ref{sec:environment}) at the redshift of each sample galaxy $i$. For each sample galaxy we count the number of \texttt{ransack} points within $\theta_{i, \rm{ap}}$ of the galaxy: $n_{i,\rm{ransack}}$. Afterwards, we compare $n_{i,\rm{ransack}}$ to the expected value computed from the angular area of the environment defining aperture and the EDP window function: 
\begin{equation} \label{eq:ransack}
<n_{\rm{ransack}}>_{i} = \frac{N_{\rm{ransack}}}{A_{\rm{EDP}}}\times {\pi \theta_{i, \rm{ap}}^2} \times f_{\rm{thresh}}. 
\end{equation} 
$A_{\rm{EDP}}$ is the total angular area of the EDP window function and $f_{\rm{thresh}}$ is the fractional threshold for the edge effect cut-off. For $R_{\rm{ap}}= 2\;h^{-1} \rm{Mpc}$, $f_{\rm{thresh}} = 0.75$. If $n_{i, \rm{ransack}}$ is greater than $<n_{\rm{ransack}}>_i$ then galaxy $i$ remains in our sample; otherwise, it is discarded.

@@

Although we imposed PRIMUS redshift errors on our SDSS galaxies in order to consistently measure environment throughout our entire sample, we note a significant discrepancy between the SDSS and PRIMUS samples in the fractional contribution of each environment classification within the redshift bin. For each of the PRIMUS redshift bins, approximately $50 \%$ of galaxies in the redshift bin are in low density environments and roughly $30 \%$ are in high density environments. In contrast, in the SDSS redshift bin, $38 \%$ of galaxies in the redshift bin are in low density environments and $45 \%$ are in high density environments. We remind the reader that this is due to the different mass-completeness limits imposed on our galaxy sample for each redshift bins. 
\begin{figure*}
  \begin{center}
    \leavevmode
    \epsfig{file=fig_smf_cylr2h25_thresh75_bin0_20_sdss_z0_05_0_12_primus_z0_2_1_0_2envbin_lowenv0_5_highenv3_0_lit_primuszerr.eps, height=0.75\textwidth,angle=0}
     \caption{Evolution of stellar mass functions of star-forming (top) and quiescent (bottom) target galaxies in 
low (left) and high (right) environments from redshift range $z=0-0.8$. The environment of each galaxy  
was calculated using a cylindrical aperture size of $R=2 \: \rm{Mpc}$ and $H=25 \: \rm{Mpc}$ and  
classification based on the cut-offs specified in Table \ref{tab:aperture}. The SMFs use mass bins of 
width $\Delta \rm{log}(\mathcal{M}/\mathcal{M}_{\odot})=0.25$. In each panel we use shades of blue 
(star-forming) and orange (quiescent) to represent the SMF at different redshift, higher redshifts being
progressively lighter.}      \label{fig:smf}
    \end{center}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% STELLAR MASS FUNCTION
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Stellar Mass Function} \label{sec:smf}
Our galaxy sample has so far been classified into quiescent or star-forming and low or high density environments. We further divide these subsamples into redshift bins: $0.05-0.12$, $0.2-0.4$, $0.4-0.6$, and $0.6-0.8$ for a total 16 subsamples. In Section \ref{sec:smfcalc}, we calculate the SMF for each of these subsamples. Then we examine the evolution of active and quiescent subsample SMFs in different environments in Section \ref{sec:smfevol}.  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% STELLAR MASS FUNCTION CALCULATIONS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Stellar Mass Function Calculations} \label{sec:smfcalc} 
To calculate the SMFs we employ a non-parametric $1/{V_{\rm{max}}}$ estimator commonly used for galaxy luminosity functions and stellar mass functions, as done in \cite{Moustakas:2013aa} and discussed in the review \cite{Johnston:2011aa}. The differential SMF is given by the following equation:
\begin{equation} \label{eq:phi}
\Phi(\rm{log}\: \mathcal{M}) \Delta(\rm{log} \:\mathcal{M}) = \sum\limits_{i=1}^{N} \frac{w_i}{V_{\rm{max,avail},i}}. 
\end{equation}
$w_i$ is the statistical weight of galaxy $i$ and $\Phi(\rm{log}\: \mathcal{M}) \Delta(\rm{log}\: \mathcal{M})$ is the number of galaxies ($N$) per unit volume within the stellar mass range $[\rm{log} \:\mathcal{M}, \rm{log} \:\mathcal{M}+\Delta(\rm{log} \:\mathcal{M})]$. The equation above is same as Equation 3. in \cite{Moustakas:2013aa} except that we use $V_{\rm{max,avail}}$ instead than $V_{\rm{max}}$, to account for the edge effects of the survey discussed in Section \ref{sec:edgeeffect}.

>>>>

Throughout the paper we assume cosmology with $\Omega_{m} = 0.3, \Omega_{\Lambda} = 0.7$, and $H_0 = 70\: \rm{km} \; \rm{s}^{-1} \rm{Mpc}^{-1}$.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% SAMPLE SELECTION
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Sample Selection} \label{sec:sample}
We are interested in quantifying the effects of galaxy environment on the evolution of the quenched fraction over the redshift range $0 < z < 1$. For our analysis, we require a sample with sufficient depth and robustness to probe the redshift range and to measure galaxy environment. PRIMUS with its $\sim 120,000$ spectroscopic redshifts provides a large data set at intermediate redshifts for our analysis. In addition, we anchor our analysis at intermediate redshifts with a low redshift sample derived from the Sloan Digital Sky Survey (\cite{York:2000aa}).

@@

For our aperture, we use a cylinder of dimensions: $R_{\rm{ap}} = 2.9 \rm{Mpc}$ and $H_{\rm{ap}} = 36 \rm{Mpc}$. Though spherical apertures are often used in literature (e.g. \citealt{Croton:2005aa}), we use a cylindrical aperture in order to account for the PRIMUS redshift errors and redshift space distortions (i.e. ``Finger of God" effect). \cite{Cooper:2005aa} finds that $\pm 1000 \; \rm{km} \; \rm{s^{-1}}$ optimally reduces the effects of redshift space distortions. The PRIMUS redshift uncertainty at $z \sim 0.7$ corresponds to $\sigma_z < 0.01$, so our aperture height choice of $36 \; \rm{Mpc}$ accounts for both of these effects. Our choice of cylinder radius was motivated by scale dependence analyses in literature (\citealt{Blanton:2006aa, Wilman:2010aa, Muldrew:2012aa}), which suggest that galactic properties such as color and quenched fractions are dependent on halo-scale properties such as host dark matter halo mass. \cite{Wilman:2010aa}, which uses environment defined by annuli of different radii, find positive correlation for quenched fraction and color on scales $< 1 \; \rm{Mpc}$. Our choice of $2.9 \rm{Mpc}$ provides sufficient sample size of galaxies in dense environments, for robust statistics, while tracing galactic properties within the halo scale.

@@

The analysis we describe below uses a fixed cylindrical aperture with dimensions $R_{\rm{ap}} = 2.9 \:\rm{Mpc}$ and $H_{\rm{ap}} = 36 \:\rm{Mpc}$ to measure environment. The same analysis was extended for varying aperture dimensions $R_{\rm{ap}} = 1.4, 2.9, 4.3 \: \rm{Mpc}$ and $H_{\rm{ap}} = 36, 71 \:\rm{Mpc}$. We also adjusted the environment classification thresholds for these analyses. The results obtained from using different apertures and environment classification thresholds are qualitatively consistent with the results presented below.

@@

To account for these edge effects, we use a Monte Carlo method to impose edge cutoffs on our galaxy sample. First, using \texttt{ransack} from \cite{Swanson:2008aa}, we construct a random sample of  $N_{\rm{ransack}} = 1,000,000$ points with $RA$ and $Dec$ randomly selected within the window function of the EDP (SDSS EDP and PRIMUS EDP separately). We then compute the angular separation, $\theta_{i, \rm{ap}}$ that corresponds to $R_{\rm{ap}}$ (Section \ref{sec:environment}) at the redshift of each sample galaxy $i$. For each sample galaxy we count the number of \texttt{ransack} points within $\theta_{i, \rm{ap}}$ of the galaxy: $n_{i,\rm{ransack}}$. Afterwards, we compare $n_{i,\rm{ransack}}$ to the expected value computed from the angular area of the environment defining aperture and the EDP window function: 
\begin{equation} \label{eq:ransack}
<n_{\rm{ransack}}>_{i} = \frac{N_{\rm{ransack}}}{A_{\rm{EDP}}}\times {\pi \theta_{i, \rm{ap}}^2} \times f_{\rm{thresh}}. 
\end{equation} 
$A_{\rm{EDP}}$ is the total angular area of the EDP window function and $f_{\rm{thresh}}$ is the fractional threshold for the edge effect cut-off. For $R_{\rm{ap}}= 2.9 \;\rm{Mpc}$, $f_{\rm{thresh}} = 0.75$. If $n_{i, \rm{ransack}}$ is greater than $<n_{\rm{ransack}}>_i$ then galaxy $i$ remains in our sample; otherwise, it is discarded.

@@

Although we imposed PRIMUS redshift errors on our SDSS galaxies in order to consistently measure environment throughout our entire sample, we note a significant discrepancy between the SDSS and PRIMUS samples in the fractional contribution of each environment classification within the redshift bin. For each of the PRIMUS redshift bins, approximately $50 \%$ of galaxies in the redshift bin are in low density environments and roughly $30 \%$ are in high density environments. In contrast, in the SDSS redshift bin, $38 \%$ of galaxies in the redshift bin are in low density environments and $45 \%$ are in high density environments. We remind the reader that this is due to the different mass-completeness limits imposed on our galaxy sample for each redshift bins. 
\begin{figure*}
  \begin{center}
    \leavevmode
    \epsfig{file=fig_smf_cylr2h25_thresh75_bin0_20_sdss_z0_05_0_12_primus_z0_2_1_0_2envbin_lowenv0_5_highenv3_0_lit_primuszerr.eps, height=0.75\textwidth,angle=0}
     \caption{Evolution of stellar mass functions of star-forming (top) and quiescent (bottom) target galaxies in 
low (left) and high (right) environments from redshift range $z=0-0.8$. The environment of each galaxy  
was calculated using a cylindrical aperture size of $R=2.9 \: \rm{Mpc}$ and $H=36 \: \rm{Mpc}$ and  
classification based on the cut-offs specified in Table \ref{tab:aperture}. The SMFs use mass bins of 
width $\Delta \rm{log}(\mathcal{M}/\mathcal{M}_{\odot})=0.25$. In each panel we use shades of blue 
(star-forming) and orange (quiescent) to represent the SMF at different redshift, higher redshifts being
progressively lighter.}      \label{fig:smf}
    \end{center}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% STELLAR MASS FUNCTION
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Stellar Mass Function} \label{sec:smf}
Our galaxy sample has so far been classified into quiescent or star-forming and low or high density environments. We further divide these subsamples into redshift bins: $0.05-0.12$, $0.2-0.4$, $0.4-0.6$, and $0.6-0.8$ for a total 16 subsamples. In Section \ref{sec:smfcalc}, we calculate the SMF for each of these subsamples. Then we examine the evolution of active and quiescent subsample SMFs in different environments in Section \ref{sec:smfevol}.  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% STELLAR MASS FUNCTION CALCULATIONS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Stellar Mass Function Calculations} \label{sec:smfcalc} 
To calculate the SMFs we employ a non-parametric $1/{V_{\rm{max}}}$ estimator commonly used for galaxy luminosity functions and stellar mass functions, as done in \cite{Moustakas:2013aa} and discussed in the review \cite{Johnston:2011aa}. The differential SMF is given by the following equation:
\begin{equation} \label{eq:phi}
\Phi(\rm{log}\: \mathcal{M}) \Delta(\rm{log} \:\mathcal{M}) = \sum\limits_{i=1}^{N} \frac{w_i}{V_{\rm{max,avail},i}}. 
\end{equation}
$w_i$ is the statistical weight of galaxy $i$ and $\Phi(\rm{log}\: \mathcal{M}) \Delta(\rm{log}\: \mathcal{M})$ is the number of galaxies ($N$) per unit volume within the stellar mass range $[\rm{log} \:\mathcal{M}, \rm{log} \:\mathcal{M}+\Delta(\rm{log} \:\mathcal{M})]$. The equation above is same as Equation 3. in \cite{Moustakas:2013aa} except that we use $V_{\rm{max,avail}}$ instead than $V_{\rm{max}}$, to account for the edge effects of the survey discussed in Section \ref{sec:edgeeffect}.
